---
title: "Strings and Factors"
author: "Lu Qiu"
date: "2023-10-17"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```

## Strings and regex

```{r}
string_vec = c("my", "name", "is", "lu")

str_detect(string_vec, "lu")
```

```{r}
str_replace(string_vec, "lu", "Lu")
```

For exact matches, you can designate matches at the beginning or end of a line.
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")
```

You can designate a list of characters that will count as a match.
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")
```

You don’t have to list these; instead, you can provide a range of letters or numbers that count as a match.
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```

The character . matches anything.
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Some characters are “special”. These include [ and ], ( and ), and .. If you want to search for these, you have to indicate they’re special using \. Unfortunately, \ is also special, so things get weird. We need to type two '\'.
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
str_detect(string_vec, '\\[[0-9]')
# special charactors
```

## Factors

```{r}
vec_sex = factor(c('male', 'male', 'female', 'female'))
vec_sex
as.numeric(vec_sex)

vec_sex = fct_relevel(vec_sex, 'male')
vec_sex
as.numeric(vec_sex)
```

## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) |> 
  html_table() |> 
  first() |>
  slice(-1)
```

need to tidy this!

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


```{r}
data_marj |>
  filter(age == "18-25") |> 
  mutate(State = fct_reorder(State, percent)) |> 
  ggplot(aes(x = State, y = percent, color = year)) + 
  geom_point() + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 
  
```

## NYC Restaurants Inspections

```{r}
data("rest_inspec")

rest_inspec |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(
    names_from = grade, 
    values_from = n) |>
  knitr::kable()
```


```{r}
rest_inspec =
  rest_inspec |>
  filter(grade %in% c("A", "B", "C"), boro != "Missing") |> 
  mutate(boro = str_to_title(boro)) |>
  mutate(dba = str_to_upper(dba))
```


```{r}
rest_inspec |> 
  filter(str_detect(dba, "PIZZA")) |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec |>
  filter(str_detect(dba, 'PIZZA')) |>
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, 'The City' = 'Manhattan')
    ) |>
  ggplot(aes(x= boro, fill = grade)) +
  geom_bar() +
  theme_bw()
```

